---
title: "Untitled"
author: "Idris"
output: html_document
---



```{r warning=FALSE}
# Loading necessary libraries
library(tidyverse)  # Data manipulation and visualization
library(caret)      # For machine learning
library(randomForest)  # For random forest model
library(xgboost)    # For XGBoost model
library(e1071)      # For SVM model
library(leaflet)    # For map visualization
library(RColorBrewer)  # For color palettes
library(viridis)    # For better color scales
library(corrplot)   # For correlation matrix
library(ggcorrplot) # For advanced correlation matrix visualization
library(ggplot2)    # For advanced plotting
```


# ==== Load Dataset ====
```{r warning=FALSE}
housing_data <- read.csv("/Users/macbook/Downloads/Housing Data_Same Region_8f72e77ef59ca0d2b5284fe2a21dbca0.csv")
head(housing_data)
```

####     GEOMAP
```{r warning=FALSE}
# Create popup content
housing_data <- housing_data %>%
  mutate(popup_info = paste0(
    "<b>Parcel:</b> ", PARCELNO, "<br>",
    "<b>Price ($):</b> ", SALE_PRC, "<br>",
    "<b>Living Area (sqft):</b> ", TOT_LVG_AREA
  ))

# Create color palette with viridis
pal <- colorNumeric(palette = "viridis", domain = housing_data$SALE_PRC)

leaflet(housing_data) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~LONGITUDE,
    lat = ~LATITUDE,
    radius = 8,              # slightly larger for better visibility
    color = ~pal(SALE_PRC),  # vibrant color by price
    stroke = TRUE,           # adds border
    weight = 1,              # border thickness
    fillOpacity = 0.9,       # more solid fill
    popup = ~popup_info
  ) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = ~SALE_PRC,
    title = "Sale Price ($)",
    opacity = 0.9
  )
```






# ==== Basic Cleaning ====
```{r warning=FALSE}
housing_data <- housing_data %>%
  select(-LATITUDE, -LONGITUDE, -PARCELNO) %>%
  select(-SALE_PRC, everything(), SALE_PRC) %>%
  distinct()
```

# ==== Convert Types ====
```{r warning=FALSE}
housing_data$structure_quality <- factor(housing_data$structure_quality, levels = c(1, 2, 3, 4, 5), ordered = TRUE)
housing_data$month_sold <- factor(housing_data$month_sold, levels = 1:12, labels = month.abb)
housing_data$avno60plus <- factor(housing_data$avno60plus, levels = c(0, 1), labels = c("No", "Yes"))
housing_data$SALE_PRC <- as.numeric(housing_data$SALE_PRC)

```

```{r warning=FALSE}

colnames(housing_data)

head(housing_data)

```




#             DESCRIPTIVE ANALYSIS  =======

```{r warning=FALSE}
# Select the numeric variables
numeric_vars <- housing_data %>% select(LND_SQFOOT, TOT_LVG_AREA, SPEC_FEAT_VAL, RAIL_DIST, OCEAN_DIST, WATER_DIST, CNTR_DIST, SUBCNTR_DI, HWY_DIST, age)

# Create a custom function to compute all the desired statistics
desc_stats <- function(x) {
  c(
    Count = length(x),
    Min = min(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE),
    Mean = mean(x, na.rm = TRUE),
    Median = median(x, na.rm = TRUE),
    Std = sd(x, na.rm = TRUE),
    Skewness = skewness(x, na.rm = TRUE),
    Kurtosis = kurtosis(x, na.rm = TRUE)
  )
}

# Apply the function to each variable and create a summary table
stats_table <- as.data.frame(t(apply(numeric_vars, 2, desc_stats)))

# View the table with row names as variables
stats_table$Variable <- rownames(stats_table)
stats_table <- stats_table %>%
  select(Variable, Count, Min, Max, Mean, Median, Std, Skewness, Kurtosis)

# Print the result
print(stats_table)
```


#               EDA    ============

### Creating histogram for LND_SQFOOT
```{r warning=FALSE}
ggplot(housing_data, aes(x = LND_SQFOOT)) +
  geom_histogram(binwidth = 900, fill = "blue", color = "black", alpha = 0.9) +
  labs(title = "Histogram of Land Square Foot", x = "Land Square Foot", y = "Frequency") +
  theme_minimal()

# Create histogram for TOT_LVG_AREA
ggplot(housing_data, aes(x = TOT_LVG_AREA)) +
  geom_histogram(binwidth = 100, fill = "green", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Total Living Area", x = "Total Living Area (sqft)", y = "Frequency") +
  theme_minimal()

# Create histogram for SALE_PRC
ggplot(housing_data, aes(x = SALE_PRC)) +
  geom_histogram(binwidth = 50000, fill = "red", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Sale Price", x = "Sale Price", y = "Frequency") +
  theme_minimal()


# Create histogram for LND_SQFOOT
ggplot(housing_data, aes(x = age)) +
  geom_histogram(bins = 20, fill = "blue", color = "black", alpha = 0.9) +
  labs(title = "Histogram of Age", x = "Age", y = "Frequency") +
  theme_minimal()

```


```{r warning=FALSE}
# Boxplot for SALE_PRC (Sale Price)
ggplot(housing_data, aes(x = "", y = SALE_PRC)) + 
  geom_boxplot(outlier.colour = "red", outlier.size = 2) + 
  labs(title = "Boxplot of SALE_PRC (Sale Price)", y = "Sale Price") +
  theme_minimal()

# Boxplot for TOT_LVG_AREA (Total Living Area)
ggplot(housing_data, aes(x = "", y = TOT_LVG_AREA)) + 
  geom_boxplot(outlier.colour = "red", outlier.size = 2) + 
  labs(title = "Boxplot of TOT_LVG_AREA (Total Living Area)", y = "Total Living Area (sq ft)") +
  theme_minimal()

# Boxplot for LND_SQFOOT (Land Square Footage)
ggplot(housing_data, aes(x = "", y = LND_SQFOOT)) + 
  geom_boxplot(outlier.colour = "red", outlier.size = 2) + 
  labs(title = "Boxplot of LND_SQFOOT (Land Square Footage)", y = "Land Square Footage (sq ft)") +
  theme_minimal()

```

```{r warning=FALSE}
## Histogram for each of the variables with individual y-axis scales
housing_data %>%
  select(LND_SQFOOT, TOT_LVG_AREA, SPEC_FEAT_VAL) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value, fill = variable)) +
  geom_histogram(binwidth = 10000, bins = 10, alpha = 0.6, position = "identity") +
  facet_wrap(~variable, scales = "free_y") +  # Set scales to free for each variable's y-axis
  labs(title = "Histogram for LND_SQFOOT, TOT_LVG_AREA, SPEC_FEAT_VAL", x = "Value", y = "Frequency") +
  theme_minimal()
```


## Boxplot for each of the variables with their respective y-axis on the same page
```{r warning=FALSE}
housing_data %>%
  select(LND_SQFOOT, TOT_LVG_AREA, SPEC_FEAT_VAL) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = 1, y = value)) +  # Fix x-axis as 1, so only one boxplot per plot
  geom_boxplot(fill = "lightblue", color = "black") +
  facet_wrap(~variable, scales = "free_y") +  # Independent y-axes for each variable
  labs(title = "Boxplot for LND_SQFOOT, TOT_LVG_AREA and SPEC_FEAT_VAL", x = "Variables", y = "Value") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())  # Remove x-axis labels
```


# Density plot for each of the variables
```{r warning=FALSE}
housing_data %>%
  select(LND_SQFOOT, TOT_LVG_AREA, SPEC_FEAT_VAL) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value, fill = variable)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Density Plot for LND_SQFOOT, TOT_LVG_AREA and SPEC_FEAT_VAL", x = "Value", y = "Density")+
  theme(legend.position = "bottom")  # Moves the legend to the bottom

```


## Correlation HeatmapPlot
```{r warning=FALSE}
# Compute the correlation matrix for the selected variables
cor_matrix <- cor(housing_data[, c("LND_SQFOOT", "TOT_LVG_AREA", "SPEC_FEAT_VAL", "RAIL_DIST", 
                                   "OCEAN_DIST", "WATER_DIST", "CNTR_DIST", "SUBCNTR_DI", 
                                   "HWY_DIST", "age", "SALE_PRC")], use = "complete.obs")
cor_matrix
# Plot the correlation heatmap
corrplot(cor_matrix, method = "color", type = "full", 
         tl.col = "black", tl.cex = 0.8, cl.lim = c(-1, 1), 
         addCoef.col = "white", number.cex = 0.7, 
         title = "Correlation Heatmap of Housing Data", 
         mar = c(0, 0, 2, 0))

```

```{r warning=FALSE}
# Visualize correlation matrix
library(ggcorrplot)
ggcorrplot(cor_matrix, lab = TRUE, type = "lower", colors = c("red", "white", "blue"))
```



# Create boxplot usingSale Price by Structure Quality
```{r warning=FALSE}
ggplot(housing_data, aes(x = structure_quality, y = SALE_PRC)) +
  geom_boxplot(fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Sale Price by Structure Quality", 
       x = "Structure Quality", 
       y = "Sale Price") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Bar chart for structure_quality with data labels
```{r warning=FALSE}
ggplot(housing_data, aes(x = structure_quality)) + 
  geom_bar(fill = "skyblue", color = "black") + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, color = "black") +  # Add data labels
  labs(title = "Distribution of Structure Quality", x = "Structure Quality", y = "Count") +
  theme_minimal()
```
# Bar chart for month_sold with data labels
```{r warning=FALSE}
ggplot(housing_data, aes(x = month_sold)) + 
  geom_bar(fill = "lightgreen", color = "black") + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, color = "black") +  # Add data labels
  labs(title = "Distribution of Month Sold", x = "Month Sold", y = "Count") +
  theme_minimal()

```

# Line chart for month_sold with data labels
```{r warning=FALSE}
ggplot(housing_data, aes(x = month_sold)) + 
  stat_count(geom = "line", aes(group = 1), color = "blue", size = 1) +  # Line chart
  geom_point(aes(y = ..count..), stat = "count", color = "red", size = 3) +  # Data points
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, color = "black") +  # Data labels
  labs(title = "Distribution of Month Sold", x = "Month Sold", y = "Count") +
  scale_y_continuous(limits = c(650, NA)) +  # Set y-axis limit starting from 650
  theme_minimal()
```


# Bar chart for avno60plus with data labels
```{r warning=FALSE}
ggplot(housing_data, aes(x = avno60plus)) + 
  geom_bar(fill = "lightcoral", color = "black") + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, color = "black") +  # Add data labels
  labs(title = "Distribution of Properties with People Over 60", x = "60+ Presence", y = "Count") +
  theme_minimal()
```
# Pie chart for avno60plus with data labels (percentage)
```{r warning=FALSE}
ggplot(housing_data, aes(x = "", fill = avno60plus)) + 
  geom_bar(width = 1, stat = "count", color = "black") + 
  coord_polar(theta = "y") +
  geom_text(stat = "count", aes(label = paste0(round((..count.. / sum(..count..)) * 100, 1), "%")), 
            position = position_stack(vjust = 0.5), color = "white") +  # Add percentage labels
  labs(title = "Pie Chart of Properties with People Over 60") +
  theme_minimal() +
  theme(axis.text.x = element_blank())  # Remove axis text for cleaner look
```



#### DISTANCE FEATURES

# Boxplot for each distance variable to check for outliers
```{r warning=FALSE}
housing_data %>%
  select(RAIL_DIST, OCEAN_DIST, WATER_DIST, CNTR_DIST, SUBCNTR_DI, HWY_DIST) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Boxplot for Distance Variables", x = "Distance Variables", y = "Distance")
```

# Histogram for each distance variable
```{r warning=FALSE}
housing_data %>%
  select(RAIL_DIST, OCEAN_DIST, WATER_DIST, CNTR_DIST, SUBCNTR_DI, HWY_DIST) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value, fill = variable)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Histogram for Distance Variables", x = "Distance", y = "Frequency")

# Density plot for each distance variable
housing_data %>%
  select(RAIL_DIST, OCEAN_DIST, WATER_DIST, CNTR_DIST, SUBCNTR_DI, HWY_DIST) %>%
  gather(key = "variable", value = "value") %>%
  ggplot(aes(x = value, fill = variable)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Density Plot for Distance Variables", x = "Distance", y = "Density")
```


# ====  Verify Data Quality
```{r warning=FALSE}

anyDuplicated(housing_data)

# Check for missing values in distance variables
sum(is.na(housing_data[, c("RAIL_DIST", "OCEAN_DIST", "WATER_DIST", "CNTR_DIST", "SUBCNTR_DI", "HWY_DIST")]))
```

# ==== Feature Engineering ====
```{r warning=FALSE}
housing_data <- housing_data %>%
  mutate(
    # Log transform for skewed numeric features
    log_LND_SQFOOT = log1p(LND_SQFOOT),
    log_TOT_LVG_AREA = log1p(TOT_LVG_AREA),
    log_SPEC_FEAT_VAL = log1p(SPEC_FEAT_VAL),
    
    # Age group bins
    age_group = cut(age, breaks = c(0, 30, 60, 100), labels = c("Young", "Middle", "Old")),
    
    # Interaction features
    area_ratio = TOT_LVG_AREA / (LND_SQFOOT + 1),
    
    # Polynomial feature
    age_sq = age^2
  )

# Convert new factor
housing_data$age_group <- factor(housing_data$age_group)


housing_data <- na.omit(housing_data)
```
# ==== Split Features and Target ====
```{r warning=FALSE}
X_data <- housing_data[, -which(names(housing_data) == "SALE_PRC")]
y_data <- housing_data$SALE_PRC

# ==== Train-Test Split ====
set.seed(123)
train_index <- createDataPartition(y_data, p = 0.8, list = FALSE)
X_train <- X_data[train_index, ]
y_train <- y_data[train_index]
X_test <- X_data[-train_index, ]
y_test <- y_data[-train_index]
```


# Bar plot for train and test sample size using ggplot in R

```{r warning=FALSE}
# First, create a data frame for plotting
train_size <- nrow(X_train)
test_size <- nrow(X_test)

sample_sizes <- data.frame(
  Sample = c("Train", "Test"),
  Size = c(train_size, test_size)
)

# Plot the bar plot
library(ggplot2)

ggplot(sample_sizes, aes(x = Sample, y = Size, fill = Sample)) + 
  geom_bar(stat = "identity", color = "black") + 
  geom_text(aes(label = Size), vjust = -0.5, color = "black") + 
  labs(title = "Train and Test Sample Sizes", x = "Sample Type", y = "Sample Size") +
  theme_minimal() +
  scale_fill_manual(values = c("lightblue", "lightgreen"))
```

# ==== Preprocessing (Scaling) ====
```{r warning=FALSE}
pre_process <- preProcess(X_train, method = c("center", "scale"))
X_train_scaled <- predict(pre_process, X_train)
X_test_scaled <- predict(pre_process, X_test)

```


# ==== Random Forest ====
```{r warning=FALSE}
rf_model <- randomForest(
  x = X_train_scaled,
  y = y_train,
  ntree = 500,
  mtry = 3,
  importance = TRUE,
  na.action = na.omit
)
rf_pred <- predict(rf_model, X_test_scaled)
rf_rmse <- RMSE(rf_pred, y_test)
rf_r2 <- R2(rf_pred, y_test)
rf_mae <- MAE(rf_pred, y_test)
rf_mse <- mean((rf_pred - y_test)^2)

# Plot Random Forest Feature Importance
importance_rf <- rf_model$importance
importance_rf_df <- data.frame(Feature = rownames(importance_rf), Importance = importance_rf[, 1])

# Plot the importance
library(ggplot2)
ggplot(importance_rf_df, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  coord_flip() + 
  labs(title = "Feature Importance in Random Forest", x = "Features", y = "Importance") +
  theme_minimal()

```

# ==== XGBoost ====
```{r warning=FALSE}
X_train_matrix <- model.matrix(~ . - 1, data = X_train_scaled)
X_test_matrix <- model.matrix(~ . - 1, data = X_test_scaled)

X_train_matrix <- xgb.DMatrix(data = X_train_matrix, label = y_train)
X_test_matrix <- xgb.DMatrix(data = X_test_matrix)

xgb_model <- xgboost(
  data = X_train_matrix,
  label = y_train,
  nrounds = 100,
  objective = "reg:squarederror",
  verbose = 0
)
xgb_pred <- predict(xgb_model, X_test_matrix)
xgb_rmse <- RMSE(xgb_pred, y_test)
xgb_r2 <- R2(xgb_pred, y_test)
xgb_mae <- MAE(xgb_pred, y_test)
xgb_mse <- mean((xgb_pred - y_test)^2)


# Plot XGBoost Feature Importance
xgb.importance(model = xgb_model) %>%
  xgb.plot.importance(top_n = 10, main = "Top 10 Features in XGBoost")
```

# ==== Linear Regression ====
```{r warning=FALSE}
lm_model <- train(SALE_PRC ~ ., data = cbind(X_train_scaled, SALE_PRC = y_train), method = "lm")
summary(lm_model)
lm_pred <- predict(lm_model, newdata = X_test_scaled)

lm_rmse <- RMSE(lm_pred, y_test)
lm_r2 <- R2(lm_pred, y_test)
lm_mae <- MAE(lm_pred, y_test)
lm_mse <- mean((lm_pred - y_test)^2)
```


# ==== SVM ====
```{r warning=FALSE}
svm_model <- train(
  SALE_PRC ~ ., 
  data = cbind(X_train_scaled, SALE_PRC = y_train), 
  method = "svmRadial",
  trControl = trainControl(method = "cv", number = 5),
  preProcess = c("center", "scale"),
  tuneLength = 5
)
svm_pred <- predict(svm_model, newdata = X_test_scaled)
svm_rmse <- RMSE(svm_pred, y_test)
svm_r2 <- R2(svm_pred, y_test)
svm_mae <- MAE(svm_pred, y_test)
svm_mse <- mean((svm_pred - y_test)^2)
```


# ==== Decision Tree ====
```{r warning=FALSE}
library(rpart)

# Train Decision Tree Model
dt_model <- rpart(
  SALE_PRC ~ ., 
  data = cbind(X_train_scaled, SALE_PRC = y_train), 
  method = "anova"
)

# Predictions for Decision Tree
dt_pred <- predict(dt_model, X_test_scaled)
dt_rmse <- RMSE(dt_pred, y_test)
dt_r2 <- R2(dt_pred, y_test)
dt_mae <- MAE(dt_pred, y_test)
dt_mse <- mean((dt_pred - y_test)^2)

library(rpart.plot)
# Plot Decision Tree
rpart.plot(dt_model, main = "Decision Tree for Predicting Sale Price")
```




# ==== Combine Evaluation Metrics ====
```{r warning=FALSE}
metrics_table <- data.frame(
  Model = c("Random Forest", "XGBoost", "Linear Regression", "SVM", "Decision Tree"),
  RMSE = c(rf_rmse, xgb_rmse, lm_rmse, svm_rmse, dt_rmse),
  R_squared = c(rf_r2, xgb_r2, lm_r2, svm_r2, dt_r2),
  MAE = c(rf_mae, xgb_mae, lm_mae, svm_mae, dt_mae),
  MSE = c(rf_mse, xgb_mse, lm_mse, svm_mse, dt_mse)
)

print(metrics_table)
```
# Combine R² values for comparison
```{r warning=FALSE}
comparison_r2_data <- data.frame(
  Model = c("Random Forest", "XGBoost", "Linear Regression", "SVM", "Decision Tree"),
  R2 = c(rf_r2, xgb_r2, lm_r2, svm_r2, dt_r2)
)

# Plot comparison of R²
ggplot(comparison_r2_data, aes(x = Model, y = R2, fill = Model)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(R2, 2)), vjust = -0.3, color = "black", size = 5) +  # Adding data labels
  labs(title = "R² Comparison Across Models", x = "Model", y = "R²") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")

```


# Create a data frame for residuals comparison
```{r warning=FALSE}
residuals_data <- data.frame(
  Actual = y_test,
  RF_Pred = rf_pred,
  XGB_Pred = xgb_pred,
  LM_Pred = lm_pred,
  SVM_Pred = svm_pred,
  DT_Pred = dt_pred
)

# Plot Residuals for Random Forest
ggplot(residuals_data, aes(x = Actual, y = RF_Pred - Actual)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Residuals Plot for Random Forest", x = "Actual Sale Price", y = "Residuals") +
  theme_minimal()

ggplot(residuals_data, aes(x = Actual, y = XGB_Pred - Actual)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Residuals Plot for XGB Model", x = "Actual Sale Price", y = "Residuals") +
  theme_minimal()

ggplot(residuals_data, aes(x = Actual, y = LM_Pred - Actual)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Residuals Plot for Linear Model", x = "Actual Sale Price", y = "Residuals") +
  theme_minimal()

ggplot(residuals_data, aes(x = Actual, y = SVM_Pred - Actual)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Residuals Plot for SVM Model", x = "Actual Sale Price", y = "Residuals") +
  theme_minimal()

# ==== Residuals Plot for Decision Tree ====
ggplot(residuals_data, aes(x = Actual, y = dt_pred - Actual)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Residuals Plot for Decision Tree Model", x = "Actual Sale Price", y = "Residuals") +
  theme_minimal()

```



# =================  PREDICTION
### ==== INPUT SETUP ====
```{r warning=FALSE}
input_data <- data.frame(
  PARCELNO = "728980145245",
  LND_SQFOOT = 11247,
  TOT_LVG_AREA = 4552,
  SPEC_FEAT_VAL = 2105,
  RAIL_DIST = 4871.9,
  OCEAN_DIST = 18507.2,
  WATER_DIST = 375.8,
  CNTR_DIST = 43897.9,
  SUBCNTR_DI = 40115.7,
  HWY_DIST = 41917.1,
  age = 42,
  avno60plus = 0,
  month_sold = 8,
  structure_quality = 5
)
```

# ==== FEATURE ENGINEERING ====
```{r warning=FALSE}
input_data <- input_data %>%
  mutate(
    # Log transform for skewed numeric features
    log_LND_SQFOOT = log1p(LND_SQFOOT),
    log_TOT_LVG_AREA = log1p(TOT_LVG_AREA),
    log_SPEC_FEAT_VAL = log1p(SPEC_FEAT_VAL),
    
    # Age group bins
    age_group = cut(age, breaks = c(0, 30, 60, 100), labels = c("Young", "Middle", "Old")),
    
    # Interaction features
    area_ratio = TOT_LVG_AREA / (LND_SQFOOT + 1),
    
    # Polynomial feature
    age_sq = age^2
  )

# ==== FACTORS ====
input_data$structure_quality <- factor(input_data$structure_quality, levels = c(1, 2, 3, 4, 5), ordered = TRUE)
input_data$month_sold <- factor(input_data$month_sold, levels = 1:12, labels = month.abb)
input_data$avno60plus <- factor(input_data$avno60plus, levels = c(0, 1), labels = c("No", "Yes"))

# Drop PARCELNO (for prediction only)
input_data_no_id <- input_data[, !(names(input_data) %in% c("PARCELNO"))]

# ==== SCALE NUMERIC FEATURES ====
# List of numeric columns to scale (exclude log-transformed and categorical columns)
numeric_cols <- c("LND_SQFOOT", "TOT_LVG_AREA", "SPEC_FEAT_VAL", "RAIL_DIST", "OCEAN_DIST", "WATER_DIST", 
                  "CNTR_DIST", "SUBCNTR_DI", "HWY_DIST", "age")

# Apply scaling only to the existing numeric columns
numeric_scaled <- predict(pre_process, input_data_no_id)
numeric_scaled
# Combine scaled numeric columns with categorical columns
input_combined <- cbind(
  as.data.frame(numeric_scaled),
  avno60plus = input_data_no_id$avno60plus,
  structure_quality = input_data_no_id$structure_quality,
  month_sold = input_data_no_id$month_sold
)

# Add the log-transformed features and engineered features back into the dataset
input_combined$log_LND_SQFOOT <- input_data_no_id$log_LND_SQFOOT
input_combined$log_TOT_LVG_AREA <- input_data_no_id$log_TOT_LVG_AREA
input_combined$log_SPEC_FEAT_VAL <- input_data_no_id$log_SPEC_FEAT_VAL
input_combined$age_group <- input_data_no_id$age_group
input_combined$area_ratio <- input_data_no_id$area_ratio
input_combined$age_sq <- input_data_no_id$age_sq

# Reorder columns to match the original input_data_no_id structure
input_combined <- input_combined[, names(input_data_no_id)]

# Verify column names
colnames(input_combined)

# ==== RANDOM FOREST PREDICTION ====
predicted_price_rf <- predict(rf_model, newdata = input_combined)
predicted_price_rf

# ==== XGBOOST PREDICTION ====
# One-hot encode using model.matrix
#input_matrix <- model.matrix(~ . - 1, data = input_combined)


input_matrix <- xgb.DMatrix(data = input_combined)

# Predict with XGBoost
predicted_price_xgb <- predict(xgb_model, newdata = input_matrix)
predicted_price_xgb


# ==== FINAL TABLE ====
prediction_table <- input_data
prediction_table$Predicted_RF_Price <- predicted_price_rf
prediction_table$Predicted_XGB_Price_Scaled <- predicted_price_xgb

# Remove PARCELNO from the prediction table
prediction_table$PARCELNO <- NULL

# ==== RESULT ====
print(prediction_table)
```

